Supporting Information

### 2018-03-30 SupportingInformation.Rmd
finished editing through model selection results for aphid ff

As in section \ref{Ex_count}, results are presented in $10 \times 10$ matrix format with rows
and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the

several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data. Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.

@@

```{r tab_ff_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
             format='pandoc',
             caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair\\label{tab_ff_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_CvM,
             format='pandoc',
             caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_ff_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_KS,
             format='pandoc',
             caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_ff_p_KS}",
             booktabs=TRUE
             )
```

```{r tab_ff_AICw_normal, echo=F, results='markup'}
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='pandoc',
             caption = "Aphid_firstflight : AICw for 'Normal' copula family for each selective location pair\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw,
             format='pandoc',
             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for each selective location pair\\label{tab_ff_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$UTdep_AICw,
             format='pandoc',
             caption = "Aphid_firstflight : model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_UTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Aphid_firstflight : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_LTmUT}",
             booktabs=TRUE
             )
```

>>>>

As in section \ref{Ex_count_Model_selection}, results are presented in $10 \times 10$ 
matrix format with rows
and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows 
that, for the large majority of location pairs, some non-normal copula 
was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the best-fitting 
(lowest AIC) copula family for each pair
of sampling locations (Table \ref{tab_ff_AICw}) to the AIC weight of the normal copula 
family for the same pair (Table \ref{tab_ff_AICw_normal}) confirms that the
degree of support for normal copulas was usually not comparable to the 
degree of support for other families. Goodness of fit tests
based on a Cramer-von Mises statistic (Table \ref{tab_ff_p_CvM}) usually failed to reject the 
hypothesis that the AIC-best copula family
was also an objectively adequate description of the data, and likewise for 
goodness of fit tests based on a Kolmogorov-Smirnov
statistic (Table \ref{tab_ff_p_KS}). In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent 
of bootstrap randomizations used for
goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

```{r tab_ff_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
             format='pandoc',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_AICw_normal, echo=F, results='markup'}
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='pandoc',
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for first flight data for the leaf-curlung plum aphid, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_CvM,
             format='pandoc',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid. Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_KS,
             format='pandoc',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid. Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_KS}",
             booktabs=TRUE
             )
```

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
the model-averaged (again, AIC based) upper-tail dependence statistic are in 
Tables \ref{tab_ff_LTdep_AICw} and \ref{tab_ff_UTdep_AICw}, respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula nul hypothesis.

```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw,
             format='pandoc',
             caption = "Model-averaged lower tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$UTdep_AICw,
             format='pandoc',
             caption = "Model-averaged upper tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_UTdep_AICw}",
             booktabs=TRUE
             )
```

The difference between the model-averaged lower-tail dependence statistic and the 
model-averaged upper-tail dependence statistic (lower minus upper), shown in 
Table \ref{tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third 
part of question 1. We note this result is opposite the analogous result for aphid abundance
data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail 
dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed 
stronger upper tail dependence than lower tail dependence.

```{r tab_ff_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTmUT}",
             booktabs=TRUE
             )
```
